Model Checking Differentially Private Properties

### 2018-01-28 applications.tex
add \lCase in algorithmic
fix privacy guarantee for geometric mechanism in application

Consider monitoring events of the types $a$ and $b$ continuously.
A sequence of event types is said to be an observation. Two
obsevations are {neighbors} if one can be obtained from the other 
by replacing event types any number of times, that is, if they are of
the same length. \bw{Later, we will consider a different definition of
neighborhood by restricting the number of replacements.} 
We would like to inquire the number of event types in
observations. For instance, the query result result is $1$ 
for observations $a \cdots a$ and $b \cdots b$ of arbitrary positive
lengths; and it is $2$ for other non-empty observations.

@@

\begin{algorithm}
  \begin{algorithmic}[1]
  \Procedure{FrequencyEstimator}{}
    \State{$f_a \leftarrow 0, 1$ with probability 
      $\frac{1}{2}, \frac{1}{2}$ respectively}
    \Comment{Initialization}
    \State{$f_b \leftarrow 0, 1$ with probability
      $\frac{1}{2}, \frac{1}{2}$ respectively}
    \Match{current event type $e$}
    \Comment{Processing}
        \Case{$a$}
          \State{$f_a \leftarrow 0, 1$ with probability
            $\frac{3}{8}, \frac{5}{8}$ respectively}
        \EndCase
        \Case{$b$}
          \State{$f_b \leftarrow 0, 1$ with probability
            $\frac{3}{8}, \frac{5}{8}$ respectively}
        \EndCase
    \EndMatch
    \Match{$f_a + f_b$}
    \Comment{Output}
      \Case{0}
        \State{$\mathit{out} \leftarrow 0, 1, 2$ with probability
          $\frac{2}{3}, \frac{1}{6}, \frac{1}{6}$ respectively}
      \EndCase
      \Case{1}
        \State{$\mathit{out} \leftarrow 0, 1, 2$ with probability
          $\frac{1}{3}, \frac{1}{3}, \frac{1}{3}$ respectively}
      \EndCase
      \Case{2}
        \State{$\mathit{out} \leftarrow 0, 1, 2$ with probability
          $\frac{1}{6}, \frac{1}{6}, \frac{2}{3}$ respectively}
      \EndCase
    \EndMatch
    \State{\Return $\mathit{out}$}
  \EndProcedure
  \end{algorithmic}
  \caption{FrequencyEstimator}
  \label{algorithm:frequency-estimator}
\end{algorithm}

\indent
\emph{Privacy analysis}.
In the processing phase, the bit $f_a$ for an event type $a$ can only
have two different distributions $U$ and $B$. If the event type is
never observed, $\Pr_U[f_a = 0] = \Pr_U[f_a = 1] = \frac{1}{2}$; if
the event type has been observed, we have $\Pr_B[f_a = 0] =
\frac{3}{8}$ and $\Pr_B[f_a = 1] = \frac{5}{8}$. Note that
$\frac{1}{2} \cdot \frac{3}{8} \leq \frac{1}{2} \leq 2 \cdot
\frac{3}{8}$ and $\frac{1}{2} \cdot \frac{5}{8} \leq \frac{1}{2} \leq
2 \cdot \frac{5}{8}$. For every $x \in \{ 0, 1 \}$, we have
$\frac{1}{2} \cdot \Pr_B[f_a = x] \leq \Pr_U[f_a = x] \leq 2 \cdot
\Pr_B[f_a = x]$. On neighboring observations, the distribution of
$f_a$ can only be changed from $U$ to $B$, values of $f_a$ is $(2,
0)$-differentially private and similarly for $f_b$. Finally, the
$1/2$-geometric mechanism gives $(1/2, 0)$-differential privacy. By
composition theorem, the steaming data analysis mechanism in
Algorithm~\ref{algorithm:frequency-estimator} is $(1,
0)$-differentially private.

@@

\noindent
\emph{Privacy Specification}.
For classical differential privacy, we are interested in the output
distribution. Algorithm~\ref{algorithm:frequency-estimator} has three
different outputs. By privacy analysis, the state $I$ satisfies the
following $\dpCTLstar$ formula:
\[
\dpriv{1}{0} \F 0 \wedge \dpriv{1}{0} \F 1 \wedge \dpriv{1}{0} \F 2
\]
For pan privacy, we moreover wish to measure privacy loss in internal
states. For observations of length $k$, the probabilities of reaching
each internal states need to be sufficiently close. The state $I$
therefore must satisfy
\[
\dpriv{1}{0} \F 0 \wedge \dpriv{1}{0} \F 1 \wedge \dpriv{1}{0} \F 2
\wedge
\bigwedge_{k \in \bbfN}
\dpriv{2}{0} \X^k 00 \wedge 
\dpriv{2}{0} \X^k 01 \wedge 
\dpriv{2}{0} \X^k 10 \wedge 
\dpriv{2}{0} \X^k 11.
\]

>>>>

Consider monitoring events of the types $a$ and $b$ continuously.
A sequence of event types is said to be an observation. Two
obsevations are {neighbors} if one can be obtained from the other 
by replacing event types any number of times, that is, if they are of
the same length. %\bw{Later, we will consider a different definition of
%neighborhood by restricting the number of replacements.} 
We would like to inquire the number of event types in
observations. For instance, the query result result is $1$ 
for observations $a \cdots a$ and $b \cdots b$ of arbitrary positive
lengths; and it is $2$ for other non-empty observations.

@@

\begin{algorithm}
  \begin{algorithmic}[1]
  \Procedure{FrequencyEstimator}{}
    \State{$f_a \leftarrow 0, 1$ with probability 
      $\frac{1}{2}, \frac{1}{2}$ respectively}
    \Comment{Initialization}
    \State{$f_b \leftarrow 0, 1$ with probability
      $\frac{1}{2}, \frac{1}{2}$ respectively}
    \Match{current event type $e$}
    \Comment{Processing}
        \lCase{$a$}{$f_a \leftarrow 0, 1$ with probability
            $\frac{3}{8}, \frac{5}{8}$ respectively}
        \lCase{$b$}{$f_b \leftarrow 0, 1$ with probability
            $\frac{3}{8}, \frac{5}{8}$ respectively}
    \EndMatch
    \Match{$f_a + f_b$}
    \Comment{Output}
      \lCase{0}{$\mathit{out} \leftarrow 0, 1, 2$ with probability
          $\frac{2}{3}, \frac{1}{6}, \frac{1}{6}$ respectively}
      \lCase{1}{$\mathit{out} \leftarrow 0, 1, 2$ with probability
          $\frac{1}{3}, \frac{1}{3}, \frac{1}{3}$ respectively}
      \lCase{2}{$\mathit{out} \leftarrow 0, 1, 2$ with probability
          $\frac{1}{6}, \frac{1}{6}, \frac{2}{3}$ respectively}
    \EndMatch
    \State{\Return $\mathit{out}$}
  \EndProcedure
  \end{algorithmic}
  \caption{FrequencyEstimator}
  \label{algorithm:frequency-estimator}
\end{algorithm}

\indent
\emph{Privacy analysis}.
In the processing phase, the bit $f_a$ for an event type $a$ can only
have two different distributions $U$ and $B$. If the event type is
never observed, $\Pr_U[f_a = 0] = \Pr_U[f_a = 1] = \frac{1}{2}$; if
the event type has been observed, we have $\Pr_B[f_a = 0] =
\frac{3}{8}$ and $\Pr_B[f_a = 1] = \frac{5}{8}$. Note that
$\frac{1}{2} \cdot \frac{3}{8} \leq \frac{1}{2} \leq 2 \cdot
\frac{3}{8}$ and $\frac{1}{2} \cdot \frac{5}{8} \leq \frac{1}{2} \leq
2 \cdot \frac{5}{8}$. For every $x \in \{ 0, 1 \}$, we have
$\frac{1}{2} \cdot \Pr_B[f_a = x] \leq \Pr_U[f_a = x] \leq 2 \cdot
\Pr_B[f_a = x]$. On neighboring observations, the distribution of
$f_a$ can only be changed from $U$ to $B$, values of $f_a$ is $(2,
0)$-differentially private and similarly for $f_b$. Finally, the
$1/2$-geometric mechanism gives $(2, 0)$-differential privacy. By
composition theorem, the steaming data analysis mechanism in
Algorithm~\ref{algorithm:frequency-estimator} is $(4,
0)$-differentially private.

@@

\noindent
\emph{Privacy Specification}.
For classical differential privacy, we are interested in the output
distribution. Algorithm~\ref{algorithm:frequency-estimator} has three
different outputs. By privacy analysis, the state $I$ satisfies the
following $\dpCTLstar$ formula:
\[
\dpriv{4}{0} \F 0 \wedge \dpriv{4}{0} \F 1 \wedge \dpriv{4}{0} \F 2
\]
For pan privacy, we moreover wish to measure privacy loss in internal
states. For observations of length $k$, the probabilities of reaching
each internal states need to be sufficiently close. The state $I$
therefore must satisfy
\[
\dpriv{4}{0} \F 0 \wedge \dpriv{4}{0} \F 1 \wedge \dpriv{4}{0} \F 2
\wedge
\bigwedge_{k \in \bbfN}
\dpriv{4}{0} \X^k 00 \wedge 
\dpriv{4}{0} \X^k 01 \wedge 
\dpriv{4}{0} \X^k 10 \wedge 
\dpriv{4}{0} \X^k 11.
\]
